---
title: "Lab 1 - Data visualization"
format: html
editor: visual
---

# Questions

## Part 1

```{r}
library(tidyverse)

```

### Question 1

```{r}
midwest |> ggplot(aes(x = popdensity)) +
  geom_histogram(binwidth = 100, fill = "lightblue4") +
  labs(
    x = "Population Density",
    y = "Count",
    title = "Population density of Midwesten Counties",
    subtitle = "Binwidth = 100"
  )


```

```{r}
df = midwest
midwest |> ggplot(aes(x = popdensity)) +
  geom_histogram(binwidth = 1000, fill = "lightblue4") +
  labs(
    x = "Population Density",
    y = "Count",
    title = "Population density of Midwesten Counties",
    subtitle = "Binwidth = 1,000"
  )
```

```{r}
df = midwest
midwest |> ggplot(aes(x = popdensity)) +
  geom_histogram(binwidth = 10000, fill = "lightblue4") +
  labs(
    x = "Population Density",
    y = "Count",
    title = "Population density of Midwesten Counties",
    subtitle = "Binwidth = 10,000"
  )
```

```{r}
df = midwest
midwest |> ggplot(aes(x = popdensity)) +
  geom_histogram(binwidth = 100000, fill = "lightblue4") +
  labs(
    x = "Population Density",
    y = "Count",
    title = "Population density of Midwesten Counties",
    subtitle = "Binwidth = 100,000"
  )
```

Starting with the graph of binwidth 100, we see that the graph is very jagged with some tiny gaps in between the smaller values that are not very meaningful; and the larger values which are not very frequent are difficult to see. The binwidth 1,000 graph improves many of these problems, it is more visually appealing and smooth, with the less frequent larger values being visible. The 10,000 binwidth graph eliminated much of the information the reader can get from graph, as most of the values are relativity small, but all get grouped in a couple bins. Lastly, the 100,000 binwidth graph is functionally useless as almost all the data is in one bin. Ultimately, the 1,000 binwidth graph is the most usable and appropriate for this data.

### Question 2

### Question 3

### Question 4

### Question 5

### Question 6

### Question 7

## Part 2

**Enough about the Midwest!**

### Question 8

### Question 9
